PyTracer: Automatically profiling numerical instabilities in Python

### 2021-11-30 paper.tex
Update the discussion section

\documentclass[11pt]{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{todonotes}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xcolor,colortbl}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{makecell}
\usepackage{ulem} % provides sout
\usepackage{minted} % Provides listing environment
\usepackage[htt]{hyphenat}
\usepackage{subcaption}
% \usepackage{subfigure}

@@

\newcommand{\Yohan}[1]{\color{green!75!black}\textbf{Yohan:} #1\color{black}\xspace}
\newcommand{\cross}[0]{\cellcolor{red!65}\ding{53}}
\newcommand{\valid}[0]{\cellcolor{green!75!black}\ding{51}}
\newcommand{\warn}[0]{\cellcolor{orange!75}?}
\newcommand{\scross}[0]{\ding{53}~}
\newcommand{\svalid}[0]{\ding{51}~}
\newcommand{\swarn}[0]{?~}
\newcommand{\pytracer}[0]{PyTracer\xspace}

@@

\begin{abstract}
Numerical stability is a crucial requirement of reliable scientific computing. However, despite the pervasiveness of Python in data science, analyzing large Python programs remains challenging due to the lack of scalable numerical analysis tools available for this language. We developed \pytracer, a profiler to quantify numerical instability in Python applications to address this issue. \pytracer automatically instruments Python code at a low cost ($\simeq3.5\times$) to produce numerical traces and visualize them interactively in a Plotly dashboard. We designed \pytracer to work with any numerical noise model, including Monte-Carlo Arithmetic, random rounding or, random data perturbations. We illustrate \pytracer's capabilities by testing the numerical stability of key functions in the SciPy and Scikit-learn Python libraries. \pytracer permitted pinpointing numerical instabilities caused by a LAPACK library function and fixing them.
% Numerical stability is a crucial requirement of reliable scientific computing. However, despite the pervasiveness of Python in data science, analyzing large Python programs remains challenging due to the lack of scalable numerical analysis tools available for this language. To address this issue, we developed \pytracer, a profiler to quantify numerical instability in Python applications. \pytracer automatically instruments Python code to produce numerical traces and visualize them interactively in a Plotly dashboard.
% \pytracer is designed to work with any type of numerical noise model, including Monte-Carlo Arithmetic, random rounding or random data perturbations. We illustrate \pytracer's capabilities by testing the numerical stability of key functions in the SciPy and Scikit-learn Python libraries. Finally, \pytracer slows down executions by only $3.5 \times$ on average.
\tristan{Add one sentence to summarize the findings} 
\end{abstract}

@@

The scientific Python ecosystem is a central component of scientific
analyses due to its rich offering of data manipulation, array programming,
numerical analysis, and visualization tools. In particular, libraries such
as NumPy~\cite{harris2020array}, SciPy~\cite{virtanen2020scipy}, scikit-learn~\cite{pedregosa2011scikit} or PyTorch~\cite{paszke2019pytorch} are used in hundreds of publications every year, providing a reference set of high-quality open-source core scientific libraries. Researchers have built numerous domain-specific data analysis pipelines from this ecosystem, leveraging Python's simplicity and expressivity.

@@

Through stability evaluations of the
SciPy and scikit-learn libraries, we demonstrate \pytracer as a practical solution to review the numerical stability of extensive code bases. The remainder of this manuscript describes related approaches, presents the design of \pytracer, and demonstrates its potential in various use cases.

@@

Several tools have been developed to assess numerical quality, divided into static and dynamic approaches. Static approaches typically analyze source code while dynamic approaches trace floating-point computations throughout executions. A detailed review of these approaches is presented in~\cite{cherubin2020tools}. 
\pytracer adopts the dynamic approach due to its scalability to large existing code bases. Dynamic approaches typically trace floating-point computations, detect instabilities in traces, and visualize summary statistics about instabilities. The remainder of this section reviews existing approaches for each of these steps.

@@

\label{sec:soa}
\subsection{Numerical tracing techniques}
Tools to trace floating-point computations for C, C++, or FORTRAN programs are numerous due to the prevalence of these languages in High-Performance Computing (HPC). 
The main tracing techniques are source-to-source, compiler-based transformations, and dynamic binary instrumentation.

@@

Dynamic Binary Instrumentation (DBI) operates directly on executables, without the need for recompilation or manual changes. Therefore, it is applicable to any programming language.
CraftHPC~\cite{lam2013dynamic} uses DBI to detect catastrophic cancellations at runtime.
Verrou~\cite{fevotte2016verrou} is a Valgrind~\cite{nethercote2007valgrind} based tool that dynamically replaces
floating-point operations with their stochastic arithmetic counterparts. FPDebug~\cite{benz2012dynamic} uses DBI to detect numerical inaccuracies by using shadow computations with higher precision.
Herbgring~\cite{sanchez2017finding} is a Valgrind-based tool to detect
numerical instabilities. It uses a shadow memory to detect precision losses by comparison with results obtained at higher precision. It is combined with symbolic computation to backtrack the root of the error.
We can note that all methods based on a high-precision oracle suppose that computing with a large number of bits is sufficient to obtain an accurate result, which is not always true (see for example, the famous Muller's sequence~\cite{bajard1996introduction}. %\tristan{it looks like your veritracer paper is the first time where the ``famous Muller's sequence" was described, is that so?}).

Being agnostic to the programming language is a serious advantage of the DBI method compared to source-to-source and compiler-bases methods. However, working at a binary level makes it difficult to access high-level information needed to debug and understand the source code logic. Conversely, the source-to-source approach provides a fine-grained control on the analyzed code, but it lacks scalability, as rewriting large codes can be a tedious task. Finally, the compiler-based approach takes advantage of the best of both by being automatic and having access to high-level information. However, like source-to-source, the compiler-based method is not suitable for analyzing closed-source libraries.

@@

Three main approaches exist to detect numerical instabilities: stochastic arithmetic, uncertainty or sensitivity analysis~\cite{helton2006survey}, and random seed analysis~\cite{click2011quality}.
%\tristan{add key references for uq/sa if any}
\pytracer focuses on stochastic arithmetic since it does not require assumptions about 
the noise model contrary to the sensitivity analysis.
Moreover, stochastic arithmetic allows evaluating the numerical quality of an implementation.
%\tristan{due to...}
Stochastic arithmetic leverages randomness to estimate numerical instabilities coming from the use of floating-point representations. The main idea is to treat round-off errors 
% \tristan{only round-off errors or all numerical errors?} 
as random variables and characterize them statistically.

The first published complete stochastic arithmetic framework was CESTAC~\cite{vignes1993stochastic}, where each floating-point computation is performed multiple times with different rounding modes: round-up or round-down. From these stochastic samples, one can then derive the number of significant digits in the computed result. Discrete Stochastic Arithmetic~\cite{vignes2004discrete} extends CESTAC with comparison operators implemented in the CADNA library.

Monte Carlo Arithmetic (MCA)~\cite{parker1997monte} uses the same principle but introduces two differences:
(i) a virtual precision parameter, allowing to simulate reduced working precision,  and (ii) different perturbation modes to introduce perturbations
on function inputs or output.
MCA replaces each floating-point number with its stochastic counterpart:
\[
inexact(x) =  x + \beta^{e_x - t}\xi
\]
where $\beta$ is the number base, $t$ the virtual precision and $\xi \in (-\frac{1}{2},\frac{1}{2})$ is a random uniform variable.
Virtual precision allows for the simulation of reduced working precision.
MCA can be applied in three modes: Random Rounding (RR), Precision Bounding (PB), and full MCA, which perturb either the output, the inputs, or both with the $inexact$ function. While the RR mode is equivalent to stochastic rounding, the PB mode allows identifying catastrophic cancellations.

Stochastic arithmetic uses the number of significant digits $s$ estimated among the sampled values to quantify the numerical errors.
A common formula to determine this number from MCA samples is the one reported in~\cite{parker1997monte}:
\begin{equation}
s = -\log_{\beta}{ \left| \dfrac{\sigma}{\mu} \right|} \label{eq:sig-digits}
\end{equation}
where $\mu$ and $\sigma$ are the sample mean and standard deviation of a variable sampled with MCA.  
Sohier et al.~\cite{sohier2018confidence} recently provided a rigorous generalization of this formula to include confidence intervals.

@@

Visualization covers a wide range of numerical simulation domains, from general-purpose numerical visualization with tools
such as VTK~\cite{schroeder2000visualizing} to topological data analysis~\cite{tierny2018topological}, including matrix~\cite{wu2008matrix} or tensor~\cite{kindlmann2006diffusion} visualization.
We review here tools dedicated to the analysis of numerical variability induced by the floating-point model. Despite the large number of existing tools  to locate unstable source code sections, few of them allow the user to have a global view of the numerical accuracy of their computations or compare several executions of the same program under various perturbations.

Veritracer~\cite{chatelain2018veritracer} was the first tool to propose an embedded visualizer of numerical stability over time for C, C++, and Fortran codes. It is an extension of Verificarlo~\cite{verificarlo} that offers contextual information about variables traced to allow the user to identify instabilities quickly.
However, it neither offers a dynamic interface nor includes information about high-level structures such as \texttt{struct} or \texttt{array}.
FPAvizual~\cite{gu2014fpavisual} visualizes the effects of floating-point arithmetic but only for educational purposes.
Anteater~\cite{faust2019anteater} proposes a visualizer for Python codes, but as mentioned previously, it only deals with native Python types, and the user needs to select variables to trace manually. However, Anteater provides an insightful reflection 
% \tristan{It's a bit vague. Do you simply mean: ``Anteater's GUI inspired the design of \pytracer's interface"?} 
around visual debugging beyond floating-point analysis, which has inspired the design of \pytracer's visualization interface.

@@

% \subsection{Pytracer workflow}
\pytracer adopts the three-step workflow shown in Figure~\ref{fig:workflow}. First, it uses monkey patching to replace the functions called by the application with a wrapper that saves their arguments and returned values in a trace file.  Second, it injects noise in the application, executes it multiple times, and aggregates the resulting trace files to compute summary statistics about each traced variable. Finally, a Plotly dashboard provides interactive visualizations highlighting unstable code sections.

@@

Python loads modules through an import mechanism triggered by the \texttt{import} that involves two objects: \texttt{finders} and \texttt{loaders}. \texttt{Finders} search for packages (or 
namespaces\footnote{Python distinguishes between \texttt{regular} and \texttt{namespace} packages.
A regular package is a directory that contains a \_\_init\_\_.py file and potentially subdirectories (sub-packages) 
that contain themselves a \texttt{\_\_init\_\_.py} file and so on recursively. 
The package hierarchy follows those of the directory. 
Namespace packages introduced in Python 3.3 (\href{https://www.python.org/dev/peps/pep-0420/}{PEP 420}) do not contain an
\_\_init\_\_.py file and allow for flexible directory structures. Hence, parts of the package can be located in zip files, on the network, or in separated directories. There are no functional differences between both types of packages.}) on the local storage --- such as standard packages installed through the pip package manager --- or on a distant server.
\texttt{Finders} do not load modules. Instead, they return a specification (\texttt{spec} object) that includes 
information on where to find and how to load modules.
\texttt{Loaders} create module objects, initialize them, and fill import-related attributes 
such as \texttt{\_\_spec\_\_}. 
\texttt{Loaders} then execute modules and populate their namespace. Finally, modules are bound to their import name in the \texttt{sys.modules} dictionary.

@@

Listings~\ref{fig:wrapper_creation} and~\ref{fig:generic_wrapper} show \pytracer's function instrumentation mechanism.
For each function to instrument, \pytracer's \texttt{loader} class creates a wrapper function, dynamically compiles it with the compile builtin function, and substitutes it for the original module function.  However, this simple technique does not apply to callable class instances, i.e., class instances that have a \texttt{\_\_call\_\_} attribute and are not of type \texttt{FunctionType}. Indeed, applying the wrapper technique would 1) modify the type of the class instance to \texttt{FunctionType}, which could cause syntax and semantic errors, and 2) mask class attributes, leading to \texttt{AttributeError} exceptions.
\pytracer overrides the \texttt{\_\_call\_\_} attribute with the wrapper function when possible to overcome this issue.
When the \texttt{\_\_call\_\_} attribute is not writable,  Pytracer does not instrument the class and returns a warning.

@@

\begin{listing}
    \centering
\begin{lstlisting}[language=Python,style=customPython,]
def generic_wrapper(self, info, *args, **kwargs):
    """Generic wrapper dumping inputs and outputs of the wrapped function.
    The id_dict dict keeps a mapping between the original function identifier
    and itself. Arguments are binded to avoid mispositioning.
    Returns the actual output of the wrapped function.

@@

    Returns:
        outputs: Outputs of the wrapped function
    """
    fid, fmodule, fname = info
    function = original_function_cache.id_dict[fid]
    bind = Binding(function, *args, **kwargs)
    stack = self.backtrace()
    time = elements()
    self.inputs(time=time,
                module_name=fmodule,
                function_name=fname,
                function=function,
                args=bind.arguments,
                backtrace=stack)
    outputs = function(*bind.args, **bind.kwargs)
    self.outputs(time=time,
                 module_name=fmodule,
                 function_name=fname,
                 function=function,
                 args=outputs,
                 backtrace=stack)
    return outputs
\end{lstlisting}
    \caption{\pytracer's generic wrapper function that dumps
    the inputs and outputs of the wrapped function.}
    \label{fig:generic_wrapper}
\end{listing}

@@

In functional programming, iterators traverse containers lazily, meaning that the next element in the sequence is only computed when the application uses it. This technique allows for the manipulation of virtually infinite sequences with finite memory. However, it implies that no complete mapping of the container returned by an iterator exists in memory since the application computes each element on the fly. Therefore, builtin iterators are not serializable which implies that \pytracer cannot save them to output traces. 
A workaround would be to convert iterators into explicit containers, however, this would increase the memory footprint significantly. \pytracer, therefore, does not instrument iterators.

@@

Once traces are available, \pytracer sequentially parses them and computes the mean, standard deviation, and number of significant digits (as in Equation~\ref{eq:sig-digits}) of all saved values. \pytracer saves these summary statistics in an HDF5 file, a hierarchical format that facilitates visual browsing. This operation critically relies on the ordering of call input and output objects in the traces. Therefore, it assumes that the application is single-threaded and that the control flow does not depend on a random state. 
% Techniques to overcome this limitation are discussed in the discussion section.

@@

Table~\ref{tab:pytracer_results_summary} summarizes
\pytracer's outcome on SciPy and scikit-learn examples.
The columns \textit{trace} and \textit{parse} report on the tracing and trace aggregation steps with three possible outcomes: green for success, red for numerical error in the application, and orange for errors due to the fuzzy noise injection
that are not considered as numerical errors, like for instance index errors due to an index shift.
% \tristan{Explain what this is, in a setence here or with a subsubsection if it requires more}. 
Overall, the tracing and parsing of SciPy and scikit-learn examples in Random Rounding mode was entirely successful for 32/40 examples. The remaining examples failed due to numerical errors in the libraries, which we discuss hereafter. The instrumentation with Full MCA is more invasive and reduced the number of successful executions to 12/40.

The column \textit{result} summarizes the precision for main outputs of SciPy and scikit-learn examples. 
We measured the precision in significant bits on the main example output by taking the element-wise mean for non-scalar values and, if the example produces many outputs, the average value among all of them. The numerical quality across all examples varies from 0 to 52, with an average of 44 significant bits. For both modes, we note that only one example failed due to a numerical error 
% \tristan{can we see that in the table? if not, don't comment on it, it is puzzling. Could you show that in the table by using the same error symbols as in Table 2?} 
(\texttt{face\_recognition} for RR and \textit{Bayesian Ridge Regression} for Full MCA) even though we underestimate this number since Full MCA mode is more intrusive than RR mode and most of the examples produce error caused by the fuzzy noise injection.

\begin{table}[]
    \centering
    \small
    \begin{subfigure}[t]{\linewidth}
    \centering
    \begin{tabular}{|ll|c|c|c|c|c|c|}
    \hline
    \multicolumn{2}{|c}{ \multirow{2}{*}{Application} } & \multicolumn{3}{|c|}{Random Rounding (RR)} & \multicolumn{3}{c|}{Full MCA} \\
    \cline{3-8}
    & & trace & parse & result & trace & parse & result \\
    \hline
    %        &               RR                 &                 MCA               \\
    %        &  numpy         &   no numpy      &      numpy      &     no numpy    \\
    %        & run   & parse  &  run   &  parse &  run   & parse  & run    & parse  \\

>>>>

\documentclass[11pt]{article}
\usepackage[margin=0.5in]{geometry}
\usepackage{todonotes}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{xcolor,colortbl}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,citecolor=blue!75!black,filecolor=black,urlcolor=blue!75!black}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage{makecell}
\usepackage{ulem} % provides sout
\usepackage{minted} % Provides listing environment
\usepackage[htt]{hyphenat}
\usepackage{subcaption}
\usepackage{graphicx,import}
\usepackage{booktabs}

% \usepackage{subfigure}

@@

\newcommand{\Yohan}[1]{\color{green!75!black}\textbf{Yohan:} #1\color{black}\xspace}
\newcommand{\cross}[0]{\cellcolor{red!65}\ding{53}}
\newcommand{\valid}[0]{\cellcolor{green!75!black}\ding{51}}
\newcommand{\warn}[0]{\cellcolor{orange!75}?}
\newcommand{\na}[0]{\cellcolor{gray!25}}
\newcommand{\s}[1]{\cellcolor{cyan!25}#1}
\newcommand{\scross}[0]{\ding{53}~}
\newcommand{\svalid}[0]{\ding{51}~}
\newcommand{\swarn}[0]{?~}
\newcommand{\pytracer}[0]{PyTracer\xspace}

\setlength{\arrayrulewidth}{0.1pt}

@@

\begin{abstract}
Numerical stability is a crucial requirement of reliable scientific computing. However, despite the pervasiveness of Python in data science, analyzing large Python programs remains challenging due to the lack of scalable numerical analysis tools available for this language. We developed \pytracer, a profiler to quantify numerical instability in Python applications. \pytracer automatically instruments Python code to produce numerical traces and visualize them interactively in a Plotly dashboard. We designed \pytracer to work with any numerical noise model including Monte-Carlo Arithmetic, random rounding or random data perturbations. We illustrate \pytracer's capabilities by testing the numerical stability of key functions in the SciPy and Scikit-learn Python libraries. Through these evaluations, we demonstrate \pytracer as a scalable, automatic, and generic framework for numerical evaluations in Python.
% Numerical stability is a crucial requirement of reliable scientific computing. However, despite the pervasiveness of Python in data science, analyzing large Python programs remains challenging due to the lack of scalable numerical analysis tools available for this language. To address this issue, we developed \pytracer, a profiler to quantify numerical instability in Python applications. \pytracer automatically instruments Python code to produce numerical traces and visualize them interactively in a Plotly dashboard.
% \pytracer is designed to work with any type of numerical noise model, including Monte-Carlo Arithmetic, random rounding or random data perturbations. We illustrate \pytracer's capabilities by testing the numerical stability of key functions in the SciPy and Scikit-learn Python libraries. Finally, \pytracer slows down executions by only $3.5 \times$ on average.
% \tristan{Add one sentence to summarize the findings} 
\end{abstract}

@@

The scientific Python ecosystem is a central component of data
analyses owing to its rich offering of data manipulation, array programming,
numerical analysis, and visualization tools. In particular, libraries such
as NumPy~\cite{harris2020array}, SciPy~\cite{virtanen2020scipy}, scikit-learn~\cite{pedregosa2011scikit} or PyTorch~\cite{paszke2019pytorch} are used in hundreds of publications every year, providing a reference set of high-quality open-source core scientific libraries. Researchers have built numerous domain-specific software tools from this ecosystem, leveraging Python's simplicity and expressivity.

@@

Through stability evaluations of the
SciPy and scikit-learn libraries, we demonstrate \pytracer as a practical solution to review the numerical stability of extensive code bases. This manuscript describes related approaches, presents the design of \pytracer, and demonstrates its potential in various use cases.

@@

Several tools have been developed to assess numerical quality, divided into static and dynamic approaches. Static approaches typically analyze source code while dynamic approaches trace floating-point computations throughout executions. A detailed review of these approaches is presented in~\cite{cherubin2020tools}. 
\pytracer adopts the dynamic approach due to its scalability to large existing code bases. Dynamic approaches typically trace floating-point computations, detect instabilities in traces, and visualize summary statistics about instabilities. This section reviews existing approaches for each of these steps.

@@

\label{sec:soa}
\subsection{Numerical tracing techniques}
Numerous tools exists to trace floating-point computations for C, C++, or FORTRAN programs due to the prevalence of these languages in High-Performance Computing (HPC). 
The main tracing techniques are source-to-source, compiler-based transformations, and dynamic binary instrumentation.

@@

Dynamic Binary Instrumentation (DBI) operates directly on executables, without the need for recompilation or manual changes. Therefore, it is applicable to any programming language.
CraftHPC~\cite{lam2013dynamic} uses DBI to detect catastrophic cancellations at runtime.
Verrou~\cite{fevotte2016verrou} is a Valgrind~\cite{nethercote2007valgrind} based tool that dynamically replaces
floating-point operations with their stochastic arithmetic counterparts. FPDebug~\cite{benz2012dynamic} uses DBI to detect numerical inaccuracies by using shadow computations with higher precision.
Herbgring~\cite{sanchez2017finding} is a Valgrind-based tool to detect
numerical instabilities. It uses a shadow memory to detect precision losses by comparison with results obtained at higher precision. It is combined with symbolic computation to backtrack the root of the error.
We can note that all methods based on a high-precision oracle suppose that computing with a large number of bits is sufficient to obtain an accurate result, which is not always true (see for example, the famous Muller's sequence~\cite{bajard1996introduction}). %\tristan{it looks like your veritracer paper is the first time where the ``famous Muller's sequence" was described, is that so?}).

Being agnostic to the programming language is a serious advantage of the DBI method compared to source-to-source and compiler-bases methods. However, working at the binary level makes it difficult to access high-level information needed to debug and understand the source-code logic. Conversely, the source-to-source approach provides a fine-grained control on the analyzed code, but it lacks scalability, as rewriting large codes can be a tedious task. Finally, the compiler-based approach takes advantage of the best of both by being automatic and having access to high-level information. However, like source-to-source, the compiler-based method is not suitable for analyzing closed-source libraries.

@@

Three main approaches exist to detect numerical instabilities: stochastic arithmetic, uncertainty or sensitivity analysis~\cite{helton2006survey}, and random seed analysis~\cite{click2011quality}.
%\tristan{add key references for uq/sa if any}
\pytracer focuses on stochastic arithmetic since it does not make any assumption about 
the noise model, in contrast to sensitivity analysis.
%\tristan{due to...}
Stochastic arithmetic leverages randomness to estimate numerical instabilities coming from the use of floating-point representations. The main idea is to treat round-off errors 
% \tristan{only round-off errors or all numerical errors?} 
as random variables and characterize them statistically.

The first published complete stochastic arithmetic framework was CESTAC~\cite{vignes1993stochastic}. In CESTAC, each floating-point computation is performed multiple times with different rounding modes: round-up or round-down. From these stochastic samples, one can then derive the number of significant digits in the computed result. Discrete Stochastic Arithmetic~\cite{vignes2004discrete} extends CESTAC with comparison operators implemented in the CADNA library.

Monte Carlo Arithmetic (MCA)~\cite{parker1997monte} uses the same principle but introduces two differences:
(i) a virtual precision parameter, allowing to simulate reduced working precision,  and (ii) different perturbation modes to introduce perturbations
on function inputs or output.
MCA replaces each floating-point number $x$ with a stochastic counterpart computed as:
\[
inexact(x) =  x + \beta^{e_x - t}\xi
\]
where $\beta$ is the number base, $e_x$ is the magnitude of $x$, $t$ the virtual precision and $\xi \in (-\frac{1}{2},\frac{1}{2})$ is a random uniform variable.
Virtual precision allows for the simulation of reduced working precision.
MCA can be applied in three modes: Random Rounding (RR), Precision Bounding (PB), and full MCA, which apply function $inexact$ to the output, the inputs, or both, respectively. While the RR mode is equivalent to stochastic rounding, the PB mode can also identify catastrophic cancellations.

Stochastic arithmetic quantifies numerical error using the number of significant digits $s$ estimated among the sampled values. A common formula to determine this number from MCA samples is the one reported in~\cite{parker1997monte}:
\begin{equation}
s = -\log_{\beta}{ \left| \dfrac{\sigma}{\mu} \right|} \label{eq:sig-digits}
\end{equation}
where $\mu$ and $\sigma$ are the sample mean and standard deviation of a variable sampled with MCA.  
Sohier et al.~\cite{sohier2018confidence} recently provided a generalization of this formula to include confidence intervals.

@@

Visualization covers a wide range of numerical simulation domains, from general-purpose numerical visualization with tools
such as VTK~\cite{schroeder2000visualizing} to topological data analysis~\cite{tierny2018topological}, including matrix~\cite{wu2008matrix} or tensor~\cite{kindlmann2006diffusion} visualization.
Despite the large number of existing tools to identify numerically unstable sections of source code, few allow the user to have a global view of the numerical precision of their computations or to compare several executions of the same program under various perturbations.

Veritracer~\cite{chatelain2018veritracer} was the first tool to propose an embedded visualizer of numerical stability over the execution of C, C++, and Fortran codes. Veritracer extends Verificarlo~\cite{verificarlo} with contextual information about traced variables, which allows the user to identify instabilities quickly.
However, Veritracer neither offers a dynamic interface nor includes information about high-level structures such as \texttt{struct} or \texttt{array}.
FPAvizual~\cite{gu2014fpavisual} visualizes the effects of floating-point arithmetic but only for educational purposes.
Anteater~\cite{faust2019anteater} proposes a visualizer for Python codes, but as mentioned previously, it only deals with native Python types, and the user needs to select variables to trace manually. Anteater provides an insightful reflection 
around visual debugging beyond floating-point analysis, which has inspired the design of \pytracer's visualization interface.

@@

% \subsection{Pytracer workflow}
\pytracer adopts the three-step workflow shown in Figure~\ref{fig:workflow}. First, \pytracer uses monkey patching to replace the functions called by the application with a wrapper that saves their arguments and returned values in a trace file.  Second, it injects noise in the application, executes it multiple times, and aggregates the resulting trace files to compute summary statistics about each traced variable. Finally, a Plotly dashboard provides interactive visualizations highlighting unstable code sections.

@@

Python loads modules through an import mechanism triggered by the \texttt{import} keyword that involves two objects: \texttt{finders} and \texttt{loaders}. \texttt{Finders} search for packages (or 
namespaces\footnote{Python distinguishes between \texttt{regular} and \texttt{namespace} packages.
A regular package is a directory that contains a \_\_init\_\_.py file and potentially subdirectories (sub-packages) 
that contain themselves a \texttt{\_\_init\_\_.py} file and so on recursively. 
The package hierarchy follows those of the directory. 
Namespace packages introduced in Python 3.3 (\href{https://www.python.org/dev/peps/pep-0420/}{PEP 420}) do not contain an
\_\_init\_\_.py file and allow for flexible directory structures. Hence, parts of the package can be located in zip files, on the network, or in separated directories. There are no functional differences between both types of packages.}) on the local storage --- such as standard packages installed through the pip package manager --- or on a distant server.
\texttt{Finders} do not load modules. Instead, they return a specification (\texttt{spec} object) that includes 
information on where to find and how to load modules.
\texttt{Loaders} create module objects, initialize them, and fill import-related attributes 
such as \texttt{\_\_spec\_\_}. 
\texttt{Loaders} then execute modules and populate their namespace. Finally, modules are bound to their import name in the \texttt{sys.modules} dictionary.

@@

Listings~\ref{fig:wrapper_creation} and~\ref{fig:generic_wrapper} show \pytracer's function instrumentation mechanism.
For each function to instrument, \pytracer's \texttt{loader} class creates a wrapper function, dynamically compiles it with the \texttt{compile} builtin function, and substitutes it for the original module function.  Although simple, this technique does not apply to callable class instances, i.e., class instances that have a \texttt{\_\_call\_\_} attribute and are not of type \texttt{FunctionType}. Indeed, applying the wrapper technique would 1) modify the type of the class instance to \texttt{FunctionType}, which could cause syntax and semantic errors, and 2) mask class attributes, leading to \texttt{AttributeError} exceptions.
\pytracer overrides the \texttt{\_\_call\_\_} attribute with the wrapper function when possible to overcome this issue.
When the \texttt{\_\_call\_\_} attribute is not writable,  Pytracer does not instrument the class and returns a warning.

@@

\begin{listing}
    \centering
\begin{lstlisting}[language=Python,style=customPython,]
def generic_wrapper(self, info, *args, **kwargs):
    """Generic wrapper saving inputs and outputs of the wrapped function.
    The id_dict dict keeps a mapping between the original function identifier
    and itself. Arguments are binded to avoid mispositioning.
    Returns the actual output of the wrapped function.

@@

    Returns:
        outputs: Outputs of the wrapped function
    """
    fid, fmodule, fname = info
    function = original_function_cache.id_dict[fid]
    bind = Binding(function, *args, **kwargs)
    stack = self.backtrace()
    time = elements()
    self.inputs(time=time,
                module_name=fmodule,
                function_name=fname,
                function=function,
                args=bind.arguments,
                backtrace=stack)
    outputs = function(*bind.args, **bind.kwargs)
    self.outputs(time=time,
                 module_name=fmodule,
                 function_name=fname,
                 function=function,
                 args=outputs,
                 backtrace=stack)
    return outputs
\end{lstlisting}
    \caption{\pytracer's generic wrapper function that saves
    the inputs and outputs of the wrapped function.}
    \label{fig:generic_wrapper}
\end{listing}

@@

In functional programming, iterators traverse containers lazily, meaning that the next element in the sequence is only computed when the application uses it. This technique allows for the manipulation of virtually infinite sequences with finite memory. However, it implies that no complete mapping of the container returned by an iterator exists in memory since the application computes each element on the fly. Therefore, iterators are not serializable which implies that \pytracer cannot save them to output traces. 
A workaround would be to convert iterators into explicit containers, however, this would increase the memory footprint significantly. \pytracer, therefore, does not instrument non-builtin iterators.

@@

Once traces are available, \pytracer sequentially parses them and computes the mean, standard deviation, and the number of significant digits (as in Equation~\ref{eq:sig-digits}) of all saved values. \pytracer saves these summary statistics in an HDF5 file, a hierarchical format that facilitates visual browsing since only
required information are loaded when requested by the user.
This operation critically relies on the ordering of call input and output objects in the traces. Therefore, it assumes that the application is single-threaded and that the control flow does not depend on a random state. 
% Techniques to overcome this limitation are discussed in the discussion section.

@@

Table~\ref{tab:pytracer_results_summary} summarizes
\pytracer's outcome on SciPy and scikit-learn examples.
Columns \textit{trace} and \textit{parse} report on the tracing and trace aggregation steps with three possible outcomes: green for success, red for numerical error in the application, and orange for errors due to the fuzzy noise injection
that are not considered numerical errors.
% , like for instance index errors due to an index shift.
% \tristan{Explain what this is, in a setence here or with a subsubsection if it requires more}. 
Overall, the tracing and parsing of SciPy and scikit-learn examples in Random Rounding mode was entirely successful for 32/40 examples. The remaining examples failed due to numerical errors in the libraries, which we discuss hereafter. The instrumentation with Full MCA is more invasive and reduced the number of successful executions to 12/40.

Column \textit{sigbits} in Table~\ref{tab:pytracer_results_summary} summarizes the precision for the main outputs of SciPy and scikit-learn examples. 
We measured the precision in significant bits on the main example output by taking the element-wise mean for non-scalar values and, if the example produces many outputs, the average value among all of them. The numerical quality across all examples varies from 0 to 52, with an average of 44 significant bits. For both modes, we note that only two examples failed due to a numerical error 
% \tristan{can we see that in the table? if not, don't comment on it, it is puzzling. Could you show that in the table by using the same error symbols as in Table 2?} 
(\texttt{face\_recognition} for RR and \textit{Bayesian Ridge Regression} for Full MCA) even though we underestimate this number since Full MCA mode is more intrusive than RR mode and most of the examples produce error caused by the fuzzy noise injection.

\begin{table}[]
    \centering
    \small
    \begin{subfigure}[t]{\linewidth}
    \centering
    \begin{tabular}{|lll|c|c|c|c|c|c|}
    \hline
    \multicolumn{3}{|c}{ \multirow{2}{*}{Application} } & \multicolumn{3}{|c|}{Random Rounding (RR)} & \multicolumn{3}{c|}{Full MCA} \\
    \cline{4-9}
    & & & trace & parse & result & trace & parse & sigbits \\
    \hline
    %        &               RR                 &                 MCA               \\
    %        &  numpy         &   no numpy      &      numpy      &     no numpy    \\
    %        & run   & parse  &  run   &  parse &  run   & parse  & run    & parse  \\
